M.Sc Proposal

### 2010-09-15 notes/proposal.tex
Work on proposal

\section{Achieved results}
The numerical solver is being implemented in MATLAB and verified against
known solution for specific problem settings.

Laplace
Diffusion-advection (upwind)
Stokes
Jacobi, RedBlack, Vanka
Coupling
Spherical
Stress-free

The solver applies iterative relaxation to solve the system, by
iterating the equations and solving repeatedly each one of them
for the corresponding variable
(Laplace equation for $\varPhi$, Stokes equation for $(\bV, P)$
and diffusion-advection equation for $C$), where
all other variables keep their values from previous iteration
(somewhat resembling Gauss-Seidel method for linear systems).

@@

@@

ghost points
precond

>>>>

\section{Achieved results}
The numerical solver is being implemented as a MATLAB package that
provides the needed tools for numerical solution of an electrophoresis problem.

The solver applies iterative relaxation to solve the system, by
iterating the equations and solving repeatedly each one of them
for the corresponding variable
(Laplace equation for $\varPhi$, Stokes equation for $(\bV, P)$
and diffusion-advection equation for $C$), where
all other variables keep their values from previous iteration
(a ``Gauss-Seidel'' method for non-linear systems).

@@

The following linear operators are implemented in spherical coordinates using regular grid:
\begin{itemize}
  \item Scalar Laplacian (for $\varPhi$ and $C$)
  \item Vector Laplacian (for $\bV$)
  \item Gradient (for $\Phi, C$ and $P$)
  \item Divergence (for $\bV$)
\end{itemize}

@@

After substitution of boundary conditions, each differential equation is transformed to a
sparse linear system of the form $\matr{A}\bx = \vect{b}$.
Due to coupling, it is not necessary to find the exact solution -- so we apply
few relaxations on each linear system, using an appropriate preconditioner $\matr{M}$
(which depends on $\matr{A}$):
\begin{eqnarray}
  \bx_{n+1} &=& \bx_{n} + \matr{M}_\matr{A}^{-1} \left(\vect{b} - \matr{A}\bx_{n}\right)
\end{eqnarray}

After the solver has converged to a solution, the stress tensor $\mathbb{T}$ on the particle
$\cal{S} = \left\{\bx : \|\bx\| = 1\right\}$ is computed:
\begin{eqnarray}
  \vect{F} = \int_{\cal{S}} \left(\mathbb{T} \cdot \bnhat \right) ds = \vec{0}.
\end{eqnarray}
Since the problem is axisymmetric, the force should be aligned with the symmetry axis:
$\vect{F} = F \bzhat$. Moreover, since the particle is stress-free in steady flow, we
have $F = 0$ for steady state velocity $\cal{U}$.

This observation enables us to find the correct $\cal{U}$ for given $\beta$, using one-dimensional root-finding algorithm on $F_\beta(\cal{U})$ function.

\subsection{Grid choice}
$C$, $\varPhi$ and $P$ are using central grid whereas staggered grid is used for $\bV$.

\subsection{Differential Equations}
Finite volume method using finite differences.
Upwind for advection.

\subsection{Preconditioners}
There are efficient sparse implementations of 
Jacobi, Red-Black Gauss-Seidel and Vanka-type.
